P2P-LLM: Collaborative LLM Inference

### 2025-05-03 docs/paper/main.tex
Pareto

A peer-to-peer network is defined as having $n$ nodes in the network. There is an LLM model with $m$ total layers which will be devided between nodes.
Each node $N$ in the network is a tuple $N(\ell, l, c, p)$ where
\begin{enumerate}
  \item $\ell _i = (x,y) : x,y \in [0,m); y > x$ is the range of layers the node is able to serve
  \item $l_{ij} \in \mathbb{R}$ is the network latency between nodes $i$ and $j$,
  \item $c_i \in \mathbb{R}$ is the computational cost of the node, measured in miliseconds
  \item $p_i \in \mathbb{R}$ is the price the nodes charges for its compute.
\end{enumerate}

@@

In order to use the network to run the LLM, a node must first find a sequence of nodes in the graph that form a complete model.
Because the LLM layers must be computed sequentially, the node must find a path through the network that ideally will compute
the LLM in the least amount of time possible and at the lowest price. This problem can be solved by viewing the network as a directed graph where weight correspond
to the computational cost of nodes. The edges of the graph correspond the compatability between the layers that each node in serving.
Specifically, every node is connected to all nodes that are serving the consecutive layer of the model. A vizualization of this directed graph is shown in Figure \ref{fig:layered_graph}.

@@

Due to this property, the P2P network graph is a directed acyclical graph (DAG). This means that there are efficient
algorithms to find the shortest path between two nodes. The first step is to perform a topological sort on the nodes in the network.
This is an ordering of the nodes such that all neighbors of a node come after the node in the ordering.
For example, a topological ordering of the DAG in Figure \ref{fig:layered_graph} is $[0,1,2,3,4,5,6,7,8]$.
The algorithm for topological sort is shown in Appendix \ref{app:topo_sort}. With the topological order, dynamic programming can be used to find the shortest path in the network.
This algorithm will iterate over the ordering, and search all the nodes neighbors for the one with the lowest weight.
The full algorithm is shown in Appendix \ref{app:shortest_path}. Most importantly, both the topological sort and path finding algorithm run in linear time
with a computational complexity of $O(|V| + |E|)$.

@@

The optimal path is well defined when weights for nodes are clearly defined, but there is not a clear way to define
weights for nodes in this case. In the P2P-LLM network there are two competing criterea that define the weight of node: computational cost and price.
The algorithm described above uses a single objective (SO), while the optimal path in the network is a multi-objective (MO) optimization problem.
There are several way to find optimal solutions to MO optimization problems \citep{1599245}. The simplest way is to convert it to a SO optimization problem, and
since the shortest path algorithm is designed for SO objective, it is the most practical solution for this task. In order to reframe the problem as a single objective,
a weighed aggreagte of the objetives is used. In the case for the peer-to-peer network, the SO is a weighed sum of computational cost $c_i$ and price $p_i$.
The objective, or weight of a node, $w_i$ is defined as.
$$w_i = \alpha c_i + 1 - \alpha p_i$$

\begin{tikzpicture}
	\begin{axis}[
			width=12cm,
			height=10cm,
			xlabel={Computational Cost},
			ylabel={Price},
			xlabel style={font=\large},
			ylabel style={font=\large},
			xmin=0, xmax=10,
			ymin=0, ymax=10,
			xtick={0,2,4,6,8,10},
			ytick={0,2,4,6,8,10},
			grid=both,
			title={Pareto Frontier: Computational Cost vs. Price},
			legend pos=north east,
			legend style={font=\footnotesize}
		]

@@

	\end{axis}
\end{tikzpicture}

@@

\subsection{Network Equilibrium}

>>>>

A peer-to-peer network is defined as having $n$ nodes in the network. There is an LLM model with $m$ total layers which will be devided between nodes.
Each node $N$ in the network is a tuple $N(\ell, l, c, p)$ where
\begin{itemize}
	\item $\ell _i = (x,y) : x,y \in [0,m); y > x$ is the range of layers the node is able to serve
	\item $l_{ij} \in \mathbb{R}$ is the network latency between nodes $i$ and $j$,
	\item $c_i \in \mathbb{R}$ is the computational cost of the node, measured in miliseconds
	\item $p_i \in \mathbb{R}$ is the price the nodes charges for its compute.
\end{itemize}

@@

In order to use the network to run the LLM, a node must first find a sequence of nodes in the graph that form a complete model.
Because the LLM layers must be computed sequentially, the node must find a path through the network that will ideally compute
the LLM in the least amount of time possible and at the lowest price. This problem can be solved by viewing the network as a directed graph where weights correspond
to the computational cost of nodes. The edges of the graph correspond the compatability between the layers that each node in serving.
Specifically, every node is connected to all nodes that are serving the consecutive layer of the model. A vizualization of this directed graph is shown in Figure \ref{fig:layered_graph}.

@@

Due to this property, the P2P network graph is a directed acyclical graph (DAG). This means that there are efficient
algorithms to find the shortest path between two nodes. The first step is to perform a topological sort on the nodes in the network.
This is an ordering of the nodes such that all neighbors of a node come after the node in the ordering.
For any graph $G = (V, E)$ a totological ordering is an order $\prec$ on $V$ such that if $(u \rightarrow v) \in E$ then $u \prec v$.
For example, a topological ordering of the DAG in Figure \ref{fig:layered_graph} is $[0,1,2,3,4,5,6,7,8]$.
The algorithm for topological sort is shown in Appendix \ref{app:topo_sort}. With the topological order, dynamic programming can be used to find the shortest path in the network.
This algorithm will iterate over the ordering, and search all the node's neighbors for the one with the lowest weight.
The full algorithm is shown in Appendix \ref{app:shortest_path}. Most importantly, both the topological sort and path finding algorithm run in linear time
with a computational complexity of $O(|V| + |E|)$.

@@

The optimal path is well defined when weights for nodes are clearly defined, but there is not a clear way to define
weights for nodes in this case. In the P2P-LLM network there are two competing criterea that define the weight of node: computational cost and price.
The algorithm described above uses a single objective (SO), while the optimal path in the network is a multi-objective (MO) optimization problem.
There are several way to find optimal solutions to MO optimization problems \citep{1599245}. The simplest way is to convert it to a SO optimization problem, and
since the shortest path algorithm is designed for SO objective, it is the most practical solution for this task. In order to reframe the problem as a single objective,
a weighed aggreagte of the objetives is used. In the case for the peer-to-peer network, the SO is a weighed sum of computational cost $c_i$ and price $p_i$.
The objective, or weight, of a node, $w_i$ is defined as.
$$w_i = \alpha \cdot c_i + (1 - \alpha) \cdot p_i$$
Howver, this assumes that the weight $\alpha$ is known a priori. In general, there are many equivilent optimal solutions in a MO optimzation problem.
These are known and Pareto-optimal solutions. They are solutions such that there is no way to improve one objective without degrading any other objectives.
A solution is said to be Pareto-optimal if there are no other solutions that dominate it. A solution $x_1$ dominates $x_2$ if
\begin{enumerate}
	\item $\forall i, f_i(x_1) \leq f_i(x_2)$
	\item $\exists i, f_i(x_1) < f_i(x_2)$
\end{enumerate}
where $f_i()$ is the utility for the $i$th objective.
The set of all Pareto-optimal solutions forms a frontier (Figure \ref{fig:Pareto-optimal}) of equivilent solutions, but there is not a clear choice on which one should be used.
In the case of the SO weighted aggreagte used in the P2P-LLM network, the parameter $\alpha$ decides where on the Pareto frontier the optimal solution lies.
Due to this, each node can independently decide that importance of price and computational cost in the optimal path through the network.

\begin{figure}
	\centering
	\begin{tikzpicture}
		\begin{axis}[
				width=12cm,
				height=8cm,
				xlabel={Computational Cost},
				ylabel={Price},
				xlabel style={font=\large},
				ylabel style={font=\large},
				xmin=0, xmax=10,
				ymin=0, ymax=10,
				xtick={0,2,4,6,8,10},
				ytick={0,2,4,6,8,10},
				grid=both,
				title={Pareto Frontier: Computational Cost vs. Price},
				legend pos=north east,
				legend style={font=\footnotesize}
			]

@@

		\end{axis}
	\end{tikzpicture}
	\caption{Pareto Frontier}
	\label{fig:Pareto-optimal}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{tikzpicture}
		\begin{axis}[
				width=14cm,
				height=8cm,
				xlabel={$p_i$},
				ylabel={Expected Revenue $R(p_i, c_i)$},
				domain=0:12,
				samples=200,
				legend pos=north east,
				grid=major,
				thick,
				every axis plot/.append style={very thick},
				title={Expected Revenue Function $R(p_i, c_i)$}
			]

			% Constant denominator
			\def\C{0.05}

			% Plot the function for different constants in numerator
			\addplot[color=blue] {(x)*(exp(-(0.5*x+0.5))/(exp(-(0.5*x+0.5)) +\C))};
			\addlegendentry{$c_i = 1$}
			\addplot[color=green] {(x)*(exp(-(0.5*x+1))/(exp(-(0.5*x+1)) +\C))};
			\addlegendentry{$c_i = 2$}
			\addplot[color=red] {(x)*(exp(-(0.5*x+2))/(exp(-(0.5*x+2)) +\C))};
			\addlegendentry{$c_i = 4$}

		\end{axis}
	\end{tikzpicture}
	\caption{Expected Revenue}
	\label{fig:exp_rev}
\end{figure}
\subsection{Network Equilibrium}
